---
title: "EML Tutorial"
author: "Haley Dunleavy"
date: "10/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Programmatic Metadata

Load in EML package
```{r}
library(EML)
```

R uses name list structure. Creating an individual element with name 'me' to then populate EML fields.
```{r}

me <- list(givenName = "Haley",
           surName = "Dunleavy")

me

```

But this is a tedious and inefficient way of filling out all the EML entries. Alternatively, one can generate this same element using helper function with syntax: `eml$elementName()`

```{r}
me <- eml$individualName(givenName = "Haley", surName = "Dunleavy")

me
```

## Minimal Valid EML

* packageId
* system
* title
* creator
* contact

Have to start with list. There isn't a helper function for the top level creation.
```{r}
doc <- list(packageId = "dataset-1",
            system = "local",
            dataset = eml$dataset(title = "A minimal value EML document", 
                                  creator = eml$creator(individualName = me), 
                                  contact = eml$contact(individualName = me))
            )

eml_validate(doc)

write_eml(eml = doc, file = "files/minimal-exaple.xml")
```


# Creating a more detailed example 

```{r}
# eml creator and contact have identical schema requirements (both fall under `responsibleParty`)
me <- eml$creator(individualName = eml$individualName(givenName = "Haley",
                                                      surName = "Dunleavy"),
                  organizationName = "Center for Ecosystem Science and Society",
                  electronicMailAddress = "hd55@nau.edu",
                  userId = list(directory = "https://orcid.org",
                                userId = "https://orcid.org/0000-0002-7383-8598"))

doc <- list(packageId = "dataset-1", 
            system = "local",
            dataset = eml$dataset(title = "A more robust valid EML dataset",
                                  creator = me,
                                  contact = me))

eml_validate(doc)
```

## Adding to the document

Can then edit the eml doc directly
```{r}
doc$dataset$abstract <- "A brief but comprehensive description of the who, what, where, when, why, and how of my dataset. At least 200 words."
```

For more complex sections. They have a grouping of functions known as the `set` helper functions. E.g. `set_methods`

```{r}
doc$dataset$methods <- set_methods("files/methods.md")
doc$dataset$methods
```

E.g. `set_coverage` can be used to set temporal and spatial coverage
```{r}
doc$dataset$coverage <- set_coverage(beginDate = 2001,
                                     endDate = 2010,
                                     geographicDescription = "Alaska, United States",
                                     westBoundingCoordinate = -179.9,
                                     eastBoundingCoordinate = -120,
                                     northBoundingCoordinate = 75,
                                     southBoundingCoordinate = 55)

eml_validate(doc)
```

## Creating data script and data file entities

Adding R script
```{r rScript}
doc$dataset$otherEntity <- eml$otherEntity(entityName = "files/datfiles_processing.R",
                                           entityDescription = "Data processing script",
                                           entityType = "application/R")

doc$dataset$otherEntity$physical <- set_physical("files/datfiles_processing.R")

```

Adding csv data table
```{r csvFile}
doc$dataset$dataTable <- eml$dataTable(entityName = "files/my-data.csv",
                                       entityDescription = "Temperature data from in-situ loggers")

doc$dataset$dataTable$physical <- set_physical("files/my-data.csv")

eml_validate(doc) #comes out false because missing attribute list
```

We need to set the attributes in data table for EML to be valid. Can do so using the `data.frame` function
```{r}
atts <- data.frame(attributeName = c("time", "temperature", "site"),
                   attributeDefinition = c("time of measurement", "measured temperature in degrees Celsius", "site identifier"),
                   unit = c(NA, "celsius", NA),
                   numberType = c(NA, "real", NA),
                   formatString = c("HH:MM:SS", NA, NA),
                   definition = c(NA, NA, "site identifier"))

atts

doc$dataset$dataTable$attributeList <- set_attributes(atts,col_classes = c("Date",
                                                                           "numeric",
                                                                           "character"))

eml_validate(doc)
str(doc$dataset$dataTable$attributeList)
```

A better option is to use shiny app to help edit attributes dataframe. Fill out table and then download table
```{r}
t <- shiny_attributes(data = read.csv("files/my-data.csv"))

```

## Save file
```{r}
write_eml(eml = doc, file = "files/more-complex-example.xml")
```

# Publishing to ADC site

Uses `dataone` and `datapack` packages. Now not only your data analysis and metadata creation are reproducible, but also your publishing.

```{r}
library(dataone)
library(datapack)
```


loading in authentication token for the ADC test data repository
```{r}
options(dataone_test_token = "copy token here")
# don't want to save this to a script, always run it to the console. 
# Also it expires every 18hrs
```

Specifying the member node rather than the whole NCEAS node. Setting up dataone client we are accessing. 
```{r}
d1c <- D1Client("STAGING", "urn:node:mnTestARCTIC")
mn <- d1c@mn
```

## Create new data package

### Set up variables and complete EML file
```{r}
dp <- new("DataPackage")
dp
```

This creates an identifier for objects (would change scheme to "doi" if submitting a final version to ADC).
```{r}
data_id <- generateIdentifier(mn, scheme = "uuid")
script_id <- generateIdentifier(mn, scheme = "uuid")
metadata_id <- generateIdentifier(mn, scheme = "uuid")
```

```{r}
doc <- read_eml("files/more-complex-example.xml")
```

Adding identifiers to metadata file
```{r}
doc$packageId <- metadata_id
doc$system <- mn@identifier
```

Now need to add a url to metadata
```{r}
# set url for csv
doc$dataset$dataTable$physical$distribution$online$url <- paste0(mn@endpoint,
                                                                 "object/",
                                                                 data_id)
# set url for script
doc$dataset$otherEntity$physical$distribution$online$url <- paste0(mn@endpoint,
                                                                   "object/",
                                                                   script_id)

eml_validate(doc)
write_eml(doc, "files/more-complex-example.xml")
```

### Add to datapack object
`id` needs to match one of the format ids on DataONE
```{r}
# Add the metadata document to the package
metadataObj <- new("DataObject",
                   id = metadata_id,
                   format ="eml://ecoinformatics.org/eml-2.1.1",
                   filename = "files/more-complex-example.xml")

dp <- addMember(dp, metadataObj)
dp
```

Adding data csv table
```{r}
# Add our data file to the package
sourceObj <- new("DataObject",
                 id = data_id,
                 format = "text/csv",
                 filename = "files/my-data.csv")

dp <- addMember(dp, sourceObj, mo = metadataObj)
dp
```


Adding Rscript
```{r}
# Add our script to the package
scriptObj <- new("DataObject",
                 id = script_id,
                 format = "application/R",
                 filename = "files/datfiles_processing.R")

dp <- addMember(dp, scriptObj, mo = metadataObj)
dp
```

Describe the provenance or the relationship between script (program) and csv files (source)
```{r}
dp <- describeWorkflow(dp, sources = sourceObj, program = scriptObj)
dp
```

### Change access
Can add colleagues to the access permissions of the program
```{r}
dp <- addAccessRule(dp,
                    subject = "http://orcid.org/0000-0003-0077-4738",
                    permission = c("read","write"),
                    getIdentifiers(dp))
```

## Add to test website

```{r}
packageId <- uploadDataPackage(d1c, dp, public = TRUE)
```

