---
title: "Tidy-Data-Tutorial"
author: "Haley Dunleavy"
date: "10/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages and data
```{r loading packages and data}
library(tidyr)
library(dplyr)
library(stringr) 
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_original)
```

To access `filter` from the `stats` package:
`stats::filter()`


NOTE: Pipe operators
`%>%` is the pipe operator. It serves as an input-output function to limit the amount of intermediate dataframes created while doing a series of steps. TL;DR. Computational efficiency and legibility!

Select columns using `select()`
```{r selecting columns}
catch_data <- catch_original %>% 
  # Can select each column
  # select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  # Or can drop columns
  select(-All, -notesRegCode)
head(catch_data)
```

```{r}
summary(catch_data)
```

Change the values in the Chinook column to numeric using `mutate()`

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)
```

Investigate using `which()` and `is.na()`.

```{r}
i <- which(is.na(catch_clean$Chinook))
catch_data[i,]
#catch_clean$Chinook[i] <- 1
```

* Use `mutate` to change the I to a 1
* Use `mutate` to coerce Chinook column to numeric

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)
```

# Tidy up dataframe

Move from wide format to the long format using `pivot_longer()`. Arguments include a vector of the columns you want to collapse, a "names" column of what you will name the collapsed column names, and a "values" column of what you will names the values of collapsed column.

```{r lengthening}
catch_long <- catch_clean %>% 
  # pivot_longer(c(Chinook, Sockeye, Coho, Pink, Chum), 
  # names_to = "Species", 
  # values_to = "Catch")
  pivot_longer(-c(Region, Year), 
               names_to = "Species",
               values_to = "Catch")

head(catch_long)
```

```{r widening}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = Species, values_from = Catch)

head(catch_wide)
```

`rename()` Catch to Catch_thousands.
`mutate()` to create a new Catch column

```{r}
catch_long <- catch_long %>% 
  # rename(Catch_thousands = Catch) %>% 
  # mutate(Catch = 1000 * Catch_thousands) %>% 
  # select(-Catch_thousands)
  mutate(Catch = Catch * 1000)


head(catch_long)
```


# Summarize data

Mean catch by region using `group_by` and `summarise`

```{r}
mean_region <- catch_long %>% 
  group_by(Region, Species) %>% 
  summarise(mean_catch = mean(Catch), 
            n_obs = n())

mean_region
```

Mean catch by region of Sockeye using `filter`
```{r}
mean_sockeye_region <- catch_long %>% 
  filter(Species == "Sockeye") %>% 
  group_by(Region, Species) %>% 
  summarise(mean_catch = mean(Catch), 
            n_obs = n())

mean_sockeye_region
```


```{r}
annual_catch <- catch_long %>% 
  group_by(Year) %>% 
  summarise(total_catch = sum(Catch)) %>% 
  arrange(total_catch)
  #arrange(desc(total_catch))

annual_catch
```

# Joining dataframes

Loading in new dataframe to then join with the catch data.
```{r load region data}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
         stringsAsFactor = FALSE) %>% 
  select(code, mgmtArea)


head(region_defs)
  
```


Joining the two dataframes by region keys into a new dataframe. Joins without a `by` argument will automatically look for common column names
```{r joining}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))


region_defs <- region_defs %>% 
  rename(Region = code)
catch_joined <- left_join(catch_long, region_defs)

head(catch_joined)
```


# Separate/Unite

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_sep <- dates_df %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", remove = FALSE)

dates_sep

dates_united <- dates_sep %>% 
  unite("Date",c("month", "day", "year"), sep = "/")

dates_united

```

## Cool manipulations
Selecting characters based on placement in text (e.g. start at character # 2 and stop at #4)
```{r}
dates_df %>% 
  mutate(three_char = substr(date, start = 2, stop = 4))

```

OR start at third to last character to the end (-3 to -1) using `str_sub` in `stringr` package.
```{r}
dates_df$date %>% 
  str_sub(start = -3, end = -1)
```

## More examples
```{r more unite/separate examples}
cities_df <- data.frame(city = c("Juneau AK", 
                                 "Sitka AK", 
                                 "Anchorage AK"),
                        stringsAsFactors = FALSE)

cities_df <- cities_df %>% 
  separate(city, into = c("City", "State"), sep = " ") %>% 
  unite(city, c("City", "State"), sep = ", ")
```

# Showcase of all the skills

```{r skills summary}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                  stringsAsFactors = FALSE)
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

mean_region <- catch_original %>%
  # select all columns but these ones
  select(-All, -notesRegCode) %>% 
  # fixing data record error and converting to numeric
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) %>% 
  # converting to long form dataframe
  pivot_longer(-c(Region, Year), names_to = "species", values_to = "catch") %>%
  # converting to appropriate units
  mutate(catch = catch*1000) %>% 
  # get mean catch by region
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch)) %>% 
  # join to include region descriptions by code
  left_join(region_defs, by = c("Region" = "code"))

head(mean_region)
```

